19Q1

### 2019-05-07 19Q1.Rmd
create drugscreen CN and GE heatmaps

```{r drug_spear_calc, warning=FALSE}
g2p_druggene_tally <- g2p %>% group_by(Drug_Gene) %>% tally() %>% drop_na()

@@

```{r corr_heatmap_function}
MakeCorrHeatmap <- function(drug_cor_res, row_names, col_title, color, legend_title) {
  # drug_cor_res$Drug <- sapply(strsplit(signif_g2p_druggene$Drug_Gene, "_", fixed = TRUE), function(x) x[1])
  # drug_cor_res$Gene <- sapply(strsplit(signif_g2p_druggene$Drug_Gene, "_", fixed = TRUE), function(x) x[2])

  # signif_grid_annot <- dcast(signif_g2p_druggene, Gene ~ Drug, value.var = "InG2P", fill = 0) %>% `rownames<-`(.[,1]) %>% select(-Gene)
  # signif_grid_annot <- signif_grid_annot[g2p_gene_order, intersect(g2p_drug_order, colnames(signif_grid))]
  # signif_grid_annot <- signif_grid_annot[complete.cases(signif_grid_annot), ]
  # signif_grid_annot[signif_grid_annot == "No"] <- ""
  # signif_grid_annot[signif_grid_annot == "Yes"] <- "*"

  col_fun <- colorRamp2(c(0, 0.05), c(color, "gray97"))
  # Legend properties list
  ht_legend_list <- list(title = legend_title, title_position = "topleft", legend_direction = "horizontal", legend_width = unit(2, "in"), border = "gray70",   at = seq(0, 0.05, by = 0.01), labels = seq(0, 0.05, by = 0.01))
  ht <- Heatmap(signif_grid,
                col = col_fun,
                na_col = "gray97",
                rect_gp = gpar(col = "gray70"),
                heatmap_legend_param = ht_legend_list,
                cluster_rows = FALSE, 
                cluster_columns = FALSE, 
                show_row_names = as.logical(row_names),
                row_names_side = "left",
                row_names_gp = gpar(fontsize = 10),
                column_names_side = "bottom",
                column_names_gp = gpar(fontsize = 10),
                show_column_names = FALSE,
                bottom_annotation = HeatmapAnnotation(
                  text = anno_text(colnames(signif_grid),
                                   rot = 45,
                                   gp = gpar(fontsize = 11),
                                   offset = unit(1, "npc"),
                                   just = "right"),
                  annotation_height = unit(0.85, "in")),
                  # annotation_height = max_text_width(colnames(signif_grid))),
                row_title = "Genes",
                row_title_gp = gpar(fontface = "bold", fontsize = 14),
                column_title = col_title, 
                column_title_gp = gpar(fontface = "bold", fontsize = 14),
                column_title_side = "bottom",
                # cell_fun = function(j, i, x, y, width, height, fill) { grid.text(signif_grid_annot[i, j], x, y, gp = gpar(fontsize = 20)) }
                )
  return(ht)
}
```

```{r drug_spear_plot}

```

>>>>

```{r drug_spear_calc, warning=FALSE, eval=FALSE}
g2p_druggene_tally <- g2p %>% group_by(Drug_Gene) %>% tally() %>% drop_na()

@@

```{r corr_heatmap_function}
MakeCorrHeatmap <- function(drug_cor_res, row_names, col_title, color_range, legend_title, metric) {
  # drug_cor_res$Drug <- sapply(strsplit(signif_g2p_druggene$Drug_Gene, "_", fixed = TRUE), function(x) x[1])
  # drug_cor_res$Gene <- sapply(strsplit(signif_g2p_druggene$Drug_Gene, "_", fixed = TRUE), function(x) x[2])
  if(toupper(metric) == "CN" | toupper(metric) == "COPY NUMBER") {
    signif_grid <- dcast(drug_cor_res, Gene ~ Drug, value.var = "CN_p", fun.aggregate = mean, fill = 0) %>% `rownames<-`(.[,1]) %>% select(-Gene)
    signif_grid <- signif_grid[g2p_gene_order, intersect(g2p_drug_order, colnames(signif_grid))]
    signif_grid <- signif_grid[complete.cases(signif_grid), ]
    signif_grid[signif_grid >= 0.05] <- ""
    signif_grid[signif_grid != ""] <- "*"

    corr_grid <- dcast(drug_cor_res, Gene ~ Drug, value.var = "CN_r", fun.aggregate = mean, fill = 0) %>% `rownames<-`(.[,1]) %>% select(-Gene)
    corr_grid <- corr_grid[g2p_gene_order, intersect(g2p_drug_order, colnames(corr_grid))]
    corr_grid <- corr_grid[complete.cases(corr_grid), ]
  }

  if(toupper(metric) == "GE" | toupper(metric) == "GENE EXPRESSION") {
    signif_grid <- dcast(drug_cor_res, Gene ~ Drug, value.var = "GE_p", fun.aggregate = mean, fill = 0) %>% `rownames<-`(.[,1]) %>% select(-Gene)
    signif_grid <- signif_grid[g2p_gene_order, intersect(g2p_drug_order, colnames(signif_grid))]
    signif_grid <- signif_grid[complete.cases(signif_grid), ]
    signif_grid[signif_grid >= 0.05] <- ""
    signif_grid[signif_grid != ""] <- "*"

    corr_grid <- dcast(drug_cor_res, Gene ~ Drug, value.var = "GE_r", fun.aggregate = mean, fill = 0) %>% `rownames<-`(.[,1]) %>% select(-Gene)
    corr_grid <- corr_grid[g2p_gene_order, intersect(g2p_drug_order, colnames(corr_grid))]
    corr_grid <- corr_grid[complete.cases(corr_grid), ]
  }

  annot_grid <- dcast(drug_cor_res, Gene ~ Drug, value.var = "InG2P", fill = 0) %>% `rownames<-`(.[,1]) %>% select(-Gene)
  annot_grid <- annot_grid[g2p_gene_order, intersect(g2p_drug_order, colnames(signif_grid))]
  annot_grid <- annot_grid[complete.cases(annot_grid), ]
  annot_grid[annot_grid == "No"] <- ""
  annot_grid[annot_grid == "Yes"] <- "*"

  col_fun <- colorRamp2(color_range, c("red3", "white", "green4"))

  # Legend properties list
  ht_legend_list <- list(title_position = "topleft", legend_direction = "horizontal", legend_width = unit(2, "in"), border = "gray70", at = seq(color_range[1], color_range[3], by = 0.2), labels = seq(color_range[1], color_range[3], by = 0.2))
  ht <- Heatmap(corr_grid,
                col = col_fun,
                name = legend_title,
                na_col = "gray97",
                rect_gp = gpar(col = "gray70"),
                heatmap_legend_param = ht_legend_list,
                cluster_rows = FALSE, 
                cluster_columns = FALSE, 
                show_row_names = as.logical(row_names),
                row_names_side = "left",
                row_names_gp = gpar(fontsize = 10),
                column_names_side = "bottom",
                column_names_gp = gpar(fontsize = 10),
                show_column_names = FALSE,
                bottom_annotation = HeatmapAnnotation(
                  text = anno_text(colnames(signif_grid),
                                   rot = 45,
                                   gp = gpar(fontsize = 11),
                                   offset = unit(1, "npc"),
                                   just = "right"),
                  annotation_height = unit(0.85, "in")),
                  # annotation_height = max_text_width(colnames(signif_grid))),
                row_title = "Genes",
                row_title_gp = gpar(fontface = "bold", fontsize = 14),
                column_title = col_title, 
                column_title_gp = gpar(fontface = "bold", fontsize = 14),
                column_title_side = "top",
                cell_fun = function(j, i, x, y, width, height, fill) { grid.text(signif_grid[i, j], x, y, gp = gpar(fontsize = 20)) }
                )
  return(ht)
}
```

```{r drug_spear_plots, eval=FALSE}
# Copy number
ccle_cor <- unfactorize(filter(g2p_cor, Dataset == "CCLE"))
ccle_cn_ht <- MakeCorrHeatmap(drug_cor_res = ccle_cor, row_names = TRUE, col_title = "CCLE", color_range = c(-0.25, 0, 0.25), legend_title = "CCLE Spearman correlation", metric = "CN")
png(filename = "./plots/manuscript/ccle_cn_heatmap.png", width = 3, height = 10, units = "in", res = 500)
draw(ccle_cn_ht, heatmap_legend_side = "top")
dev.off()

ctrp_cor <- unfactorize(filter(g2p_cor, Dataset == "CTRP"))
ctrp_cn_ht <- MakeCorrHeatmap(drug_cor_res = ctrp_cor, row_names = TRUE, col_title = "CTRP", color_range = c(-0.35, 0, 0.35), legend_title = "CTRP Spearman correlation", metric = "CN")
png(filename = "./plots/manuscript/ctrp_cn_heatmap.png", width = 9.5, height = 10, units = "in", res = 500)
draw(ctrp_cn_ht, heatmap_legend_side = "top")
dev.off()

gdsc_cor <- unfactorize(filter(g2p_cor, Dataset == "GDSC"))
gdsc_cn_ht <- MakeCorrHeatmap(drug_cor_res = gdsc_cor, row_names = TRUE, col_title = "GDSC", color_range = c(-0.25, 0, 0.25), legend_title = "GDSC Spearman correlation", metric = "CN")
png(filename = "./plots/manuscript/gdsc_cn_heatmap.png", width = 5, height = 10, units = "in", res = 500)
draw(gdsc_cn_ht, heatmap_legend_side = "top")
dev.off()

ccle_cn_ht2 <- MakeCorrHeatmap(drug_cor_res = ccle_cor, row_names = TRUE, col_title = "CCLE", color_range = c(-0.35, 0, 0.35), legend_title = "Copy num. Spearman correlation", metric = "CN")
ctrp_cn_ht2 <- MakeCorrHeatmap(drug_cor_res = ctrp_cor, row_names = FALSE, col_title = "CTRP", color_range = c(-0.35, 0, 0.35), legend_title = "Copy num. Spearman correlation", metric = "CN")
gdsc_cn_ht2 <- MakeCorrHeatmap(drug_cor_res = gdsc_cor, row_names = FALSE, col_title = "GDSC", color_range = c(-0.35, 0, 0.35), legend_title = "Copy num. Spearman correlation", metric = "CN")
png(filename = "./plots/manuscript/drug_cn_heatmap.png", width = 13, height = 10, units = "in", res = 500)
draw(ccle_cn_ht2 + ctrp_cn_ht2 + gdsc_cn_ht2, gap = unit(0.25, "in"), heatmap_legend_side = "top")
dev.off()

# Gene expression
ccle_ge_ht <- MakeCorrHeatmap(drug_cor_res = ccle_cor, row_names = TRUE, col_title = "CCLE", color_range = c(-0.5, 0, 0.5), legend_title = "CCLE Spearman correlation", metric = "GE")
png(filename = "./plots/manuscript/ccle_ge_heatmap.png", width = 3, height = 10, units = "in", res = 500)
draw(ccle_ge_ht, heatmap_legend_side = "top")
dev.off()

ctrp_ge_ht <- MakeCorrHeatmap(drug_cor_res = ctrp_cor, row_names = TRUE, col_title = "CTRP", color_range = c(-0.5, 0, 0.5), legend_title = "CTRP Spearman correlation", metric = "GE")
png(filename = "./plots/manuscript/ctrp_ge_heatmap.png", width = 9.5, height = 10, units = "in", res = 500)
draw(ctrp_ge_ht, heatmap_legend_side = "top")
dev.off()

gdsc_ge_ht <- MakeCorrHeatmap(drug_cor_res = gdsc_cor, row_names = TRUE, col_title = "GDSC", color_range = c(-0.5, 0, 0.5), legend_title = "GDSC Spearman correlation", metric = "GE")
png(filename = "./plots/manuscript/gdsc_ge_heatmap.png", width = 5, height = 10, units = "in", res = 500)
draw(gdsc_ge_ht, heatmap_legend_side = "top")
dev.off()

ccle_ge_ht2 <- MakeCorrHeatmap(drug_cor_res = ccle_cor, row_names = TRUE, col_title = "CCLE", color_range = c(-0.5, 0, 0.5), legend_title = "Gene expr. Spearman correlation", metric = "GE")
ctrp_ge_ht2 <- MakeCorrHeatmap(druwg_cor_res = ctrp_cor, row_names = FALSE, col_title = "CTRP", color_range = c(-0.5, 0, 0.5), legend_title = "Gene expr. Spearman correlation", metric = "GE")
gdsc_ge_ht2 <- MakeCorrHeatmap(drug_cor_res = gdsc_cor, row_names = FALSE, col_title = "GDSC", color_range = c(-0.5, 0, 0.5), legend_title = "Gene expr. Spearman correlation", metric = "GE")
png(filename = "./plots/manuscript/drug_ge_heatmap.png", width = 13, height = 10, units = "in", res = 500)
draw(ccle_ge_ht2 + ctrp_ge_ht2 + gdsc_ge_ht2, gap = unit(0.25, "in"), heatmap_legend_side = "top")
dev.off()
```
